---
title: "EDA"
output: html_document
---

# Exploratory data analysis {.tabset}

```{r load data}
path_files <-   "../../data/selected_data/train"
train_files <- list.files(path_files)
data_names <- gsub('.{4}$', '', train_files)

for (file in train_files) {
  
  path_file <- paste0(path_files,"/",file) 
  df_name <- substr(file,1,nchar(file)-4)
  assign( df_name ,readRDS(path_file) )
  
}


```



## Simple dataset 

```{r table with basic stats simple}

  data1 %>%
  summarise( across(-(1:5), list(mean = mean), .names = "{.col}") ) %>% 
  rbind(.,
        data1 %>%
          summarise( across(-(1:5), list(mean = sd), .names = "{.col}") )
        ) %>% 
  # cbind(
  #   list(stat = c("mean","standard_dev")) %>% as.data.frame(.),
  #   .) %>% 
  `rownames<-`(c("mean","sd")) %>% 
  #print()
  #knitr::kable() -> a
  #scroll_box(width = "100%")
  
  rmarkdown::paged_table() 

```



```{r constant sensors}
    
zero_sd_selector <- data1 %>%
  summarise( across(-(1:5), list(mean = sd), .names = "{.col}")) %>% 
  as.numeric() == 0 
# sensors which have 0 sd
constant_sensors <- data1 %>% 
  select(-c(1:5)) %>% 
  .[zero_sd_selector] %>% 
  names() 
saveRDS(constant_sensors,file = here("objects/constant_sensors_data1.rds"))
sensors_all <- colnames(data1) %>% str_subset("sensor.*")
# set difference
sensors <- setdiff(sensors_all,constant_sensors)

```



As we can see sensors with the following numbers: `r str_extract(constant_sensors, "[0-9]+")`
have zero standard deviation which means that they are constant. We will disregard
them in further analysis as they carry no information.

```{r sensors plot list simple}
l_plt <- list()
for (i in seq_along(sensors)) {
  
sensor_to_plot <- sensors[i]
no_of_units_to_draw <- 10
units_to_plot <- sample(data1$unit_no %>% unique(),no_of_units_to_draw)

l_plt[[i]] <- data1 %>% 
  select(-constant_sensors) %>% 
  pivot_longer(cols = sensors, names_to = "sensor", values_to = "value") %>% 
  filter(unit_no %in% units_to_plot & sensor == sensor_to_plot) %>% 
  ggplot(aes(x=time, y=value, color = factor(unit_no)))+
  geom_line(show.legend = F)+
  ggtitle(sensor_to_plot)

  
}

```



### Plots of all sensors vs. RUL {.tabset}
```{r tabs with plots, results='asis'}

for (i in seq_along(sensors)) {
  tmp <- l_plt[[i]]
  cat("####",sensors[i],"\n")
  print(tmp)
  cat("\n\n")
}

```

## Complicated dataset
 


```{r table with basic stats complex}

  data4 %>% group_by(setting) %>% 
  summarise( across(-(1:5), list(mean = mean), .names = "{.col}") ) %>% 
  rbind(.,
        data4 %>% group_by(setting) %>% 
          summarise( across(-(1:5), list(mean = sd), .names = "{.col}") )
        ) %>% 
  # cbind(
  #   list(stat = c("mean","standard_dev")) %>% as.data.frame(.),
  #   .) %>% 
  `rownames<-`(c(paste("mean for setting",1:6),paste("sd",1:6))) %>% 
  #print()
  #knitr::kable() -> a
  #scroll_box(width = "100%")
  
  rmarkdown::paged_table(options = list(rows.print = 12)) 

```



```{r constant sensors complex}
    
zero_sd_selector <- data4 %>% group_by(setting) %>% 
  summarise( across(-(1:5), sd, .names = "{.col}")) %>% 
  abs() %>% 
  colSums() %>% 
  as.numeric() == 0 
zero_sd_selector <- zero_sd_selector[-1]
# sensors which have 0 sd
constant_sensors <- data4 %>% 
  select(-c(1:5) & ! matches("setting")) %>% 
  .[zero_sd_selector] %>% 
  names() 
saveRDS(constant_sensors,file = here("objects/constant_sensors_data4.rds"))
sensors_all <- colnames(data4) %>% str_subset("sensor.*")
# set difference
sensors <- setdiff(sensors_all,constant_sensors)

```



As we can see sensors with the following numbers: `r str_extract(constant_sensors, "[0-9]+")`
have zero standard deviation which means that they are constant. We will disregard
them in further analysis as they carry no information.

### Plots of all sensors vs. RUL {.tabset}

```{r sensors plot list complex}
l_plt_c <- list()
for (i in seq_along(sensors)) {
  
sensor_to_plot <- sensors[i]
no_of_units_to_draw <- 10
units_to_plot <- sample(data4$unit_no %>% unique(),no_of_units_to_draw)
setting_to_plot <- data4$setting %>% unique() %>%  sample(1)

l_plt_c[[i]] <- data4 %>% 
  select(-constant_sensors) %>% 
  pivot_longer(cols = sensors, names_to = "sensor", values_to = "value") %>% 
  filter(unit_no %in% units_to_plot & sensor == sensor_to_plot & setting == setting_to_plot ) %>% 
  ggplot(aes(x=time, y=value, color = factor(unit_no)))+
  geom_line(show.legend = F)+
  ggtitle(paste(sensor_to_plot,"setting:",setting_to_plot))

  
}

```



```{r tabs with plots complex, results='asis'}

for (i in seq_along(sensors)) {
  tmp <- l_plt_c[[i]]
  cat("####",sensors[i],"\n")
  print(tmp)
  cat("\n\n")
}

```

We can see that in both datasets the RUL-sensors correspondence bears a lot of nonlinearity.
We can also observe that sensors in the complicated dataset are less homogenous which may
be caused by the two different faults present in the data as well as different operational settings.

### Target variable (RUL) {.tabset}

We may finally examine the variable we want to predict:

#### Simple dataset

```{r target simple}
# maximum RUL for simple ds
m_s <- data1 %>%
  group_by(unit_no) %>% 
  summarise(max_RUL = max(RUL))%>% 
  ggplot() +
  geom_histogram(aes(x = max_RUL), bins = 20, color = "white", fill = "steelblue")+
  ggtitle("Distribution of maximal RUL")+
  xlab("maximal RUL")+
  theme(panel.background  = NULL)

# all RUL simple
a_s <- data1 %>%
  ggplot() +
  geom_histogram(aes(x = RUL), bins = 20, color = "white", fill = "steelblue")+
  ggtitle("Distribution of all RUL")+ 
  theme(panel.background  = NULL)+
  ylab("")

ggarrange(m_s,a_s)

```

#### Complicated dataset

```{r target Complicated}
# maximum RUL for simple ds
m_c <- data4 %>%
  group_by(unit_no,setting) %>% 
  summarise(max_RUL = max(RUL))%>% 
  ggplot() +
  geom_histogram(aes(x = max_RUL), bins = 20, color = "white", fill = "steelblue")+
  facet_grid(.~ setting)+
  ggtitle("Distribution of maximal RUL", subtitle = "by setting")+
  xlab("maximal RUL")+
  theme(panel.background  = NULL)


m_c
```

